---
title: "Institutional Trust as a Buffer Against Perceived Immigration Threats"
subtitle: "A Comparative Analysis of Four East Asian Societies"
author: "Zhiyang(Chris) Cheng, Jinwen(Wendy) Xiao"
format: 
  pdf:
    mainfont: "Times New Roman"
    fontsize: 12pt
    linestretch: 1.0
---


```{r, include=FALSE}
library(pacman)
p_load(tidyverse,ggplot2,haven,interactions,sjPlot,psych,patchwork,stargazer)
dta <- readRDS("data/WVS_Cross-National_Wave_7_rds_v6_0.rds")

dta$B_COUNTRY_num <- as.numeric(dta$B_COUNTRY)

#extract needed variables
df_analysis <- dta %>%
  filter(B_COUNTRY_num %in% c(344, 702, 158, 156)) %>% 
  select(
    Country = B_COUNTRY_num,
    Q121, # The Overall impact of immigrants on the country (1-5, 5=Very Good)
    Q124, Q126, Q128, Q129, # Perceived harms (Crime, Terror, Jobs, Conflict)(0-2,2 means agree, which is the most negative view)
    Q69, Q70, Q71, Q74, # Government Trust (Police, Courts, Gov, Civil Service)(1-4,1 means trust the most)
#   Q21,  #This is causing post-treatment effect. Neighbor Preference (Immigrants/Foreign workers)(1 is dislike, 2 is did not meantion)
    Q275, Q288, Q262 # Controls: Education(1-8,larger better), Income(1-10, larger better), Age
  ) %>%
  zap_labels() %>%
  mutate(across(everything(), as.numeric))

# reorgnize the variables
df_clean <- df_analysis %>%
  # Convert unusual values to R's NA
  mutate(across(everything(), ~ifelse(.x < 0, NA, .x))) %>%
  mutate(
    Region = factor(Country, 
                    levels = c(344, 702, 158, 156), 
                    labels = c("Hong Kong", "Singapore", "Taiwan", "China")),
    
    # Dependent Variable: General View (High = Positive)
    ImpactOnCountry = Q121, 
    
    # Independent Variable: Multi-dimensional Harm Index (0-2)
    X_Harm_Index = rowMeans(across(c(Q124, Q126, Q128, Q129)), na.rm = TRUE),
    
    # Moderator: Institutional Trust Index (Combined Q69, Q70, Q71, Q74)
    # Recoded: 5 - original score so that 4 = High Trust, 1 = No Trust
    Inst_Trust_Index = rowMeans(across(c(Q69, Q70, Q71, Q74), ~ 5 - .x), na.rm = TRUE),
    
    # Dummy Variable: Personal Neighbor Preference (Q21)
    # This is causing post-treatment effect and thus desert here. 
    # 1 = Mentioned (Dislikes immigrant neighbors), 0 = Not mentioned (Accepts/Neutral)
    # Neighbor_Dislike = ifelse(Q21 == 1, 1, 0),
    
    Edu = Q275, Income = Q288, Age = Q262
  ) %>%
  # Remove cases with missing values in key analytical variables
  drop_na(ImpactOnCountry, X_Harm_Index, Inst_Trust_Index, Edu)

# Mean-Centering for better analysis
df_clean <- df_clean %>%
  mutate(
    PerceivedHarm = X_Harm_Index - mean(X_Harm_Index, na.rm = TRUE),
    InstitutionTrust = Inst_Trust_Index - mean(Inst_Trust_Index, na.rm = TRUE)
  )

#model_1, simpliest
model_simpliest <- lm(ImpactOnCountry ~ InstitutionTrust + Region + Edu + Income + Age, data = df_clean)

#model_2, regression in each country with interation term
model_hkg <- lm(ImpactOnCountry ~ InstitutionTrust  + Edu + Income + Age, data = filter(df_clean, Region == "Hong Kong"))
model_sgp <- lm(ImpactOnCountry ~ InstitutionTrust  + Edu + Income + Age, data = filter(df_clean, Region == "Singapore"))
model_twn <- lm(ImpactOnCountry ~ InstitutionTrust  + Edu + Income + Age, data = filter(df_clean, Region == "Taiwan"))
model_chn <- lm(ImpactOnCountry ~ InstitutionTrust  + Edu + Income + Age, data = filter(df_clean, Region == "China"))

#model_3, regression with interation term and compare between countires
model_full <- lm(ImpactOnCountry ~ InstitutionTrust * Region + Edu + Income + Age, data = df_clean)
```

### **Research Question**

How Government Trust Shapes Views on Immigration - Evidence from East Asia

### **Summary**

Using World Values Survey data from China, Hong Kong, Taiwan, and Singapore, this study finds that viewing immigrants negatively reduces public support for immigration. At the same time, increased trust in government is associated with more positive attitudes toward immigration. In high-migration areas such as Hong Kong and Singapore, government trust clearly mitigates the negative impact of perceived immigrant harm: individuals who are concerned about immigrants but trust the state nevertheless regard immigration as less harmful. This buffering effect does not exist in mainland China or Taiwan, where daily contact with immigrants is more limited, and sentiments appear to be influenced by broad ideological convictions rather than trust in government administration.

### **Background**

This project investigates the factors influencing public attitudes toward immigrants. As immigrants have become a vital part of developed nations, they deeply affect local economic structures, cultural integration, and community welfare. Consequently, public sentiment toward immigrants can significantly shape immigration policy and influence government responses to global migration. Therefore, understanding the drivers of these attitudes is crucial.

In this research, we will focus on East Asia especially in China, Hong Kong, Taiwan, and Singapore, which share similar cultural and positional backgrounds but have different economical and political landscapes. Mainland China has a very small share of long-term foreign residents, and its policies are cautious, so most foreigners are short-term visitors or workers. Hong Kong is a migration city with many new arrivals from mainland China and a very large number of foreign domestic workers from places like the Philippines and Indonesia. Taiwan relies heavily on migrant workers from Southeast Asia and China in factories and long-term care, and also has many “new resident” families formed through cross-border marriage. Singapore is the most migration-dependent: a large share of its population is foreign workers and their families, with strict but well-developed systems that both attract high-skilled professionals and manage large numbers of low-skilled migrant workers.

We aim to investigate how the views about immigration’s impact on national development are shaped by both attitudes toward immigrants and the trust toward their Government. 

Our core idea is that trust in government can buffer the negative effect of hostile immigration attitudes on whether immigration benefits the country. People who strongly dislike immigrants are usually more pessimistic about immigration’s impact. However, if they also have high trust in the government, they may still believe that good policies can handle potential problems and turn immigration into benefits.  

Above all, we propose following assumptions:

1.  Higher levels of institutional trust are directly associated with a more positive evaluation of immigration's overall impact on the nation.

2. 对不同地区的假设

### **Data**

Data for this study come from individual-level responses in the World Values Survey (WVS), collected between 2017 and 2022. The WVS is a cross-national survey of adult populations that asks standardized questions on values, political attitudes, and social views, including items on immigration, government trust, and perceptions of national development. The analysis focuses on four East Asian settings identified by Hong Kong, Singapore, Taiwan, China.

The main outcome variable is Q121, which measures the perceived overall impact of immigrants on the country on a 1–5 scale (1 = very bad, 5 = very good). Institutional trust is measured using four questions on confidence in the police, courts, government, and civil service (Q69, Q70, Q71, Q74), originally coded from 1 (“trust completely”) to 4 (“do not trust at all”) and then reverse-coded so that higher values indicate greater trust. Education (Q275), income (Q288), and age (Q262) are included as control variables, with higher values indicating higher levels of education, income, and age.

Prior to analysis, the sample is restricted to respondents from the four selected settings, all special missing codes and negative values are recorded as NA, and cases with missing data on key variables are dropped. For models with interaction terms, both the Perceived Harm index and the Institutional Trust index are mean-centered to facilitate interpretation of coefficients. 

A dummy variable based on Q21 (whether immigrants/foreign workers are mentioned as undesirable neighbors) is not included in the main models because it is likely post-treatment and could bias estimates of the relationship between perceived harm, institutional trust, and evaluations of immigration’s impact.


```{r}
#| label: fig-trust-dist
#| fig-cap: "Institutional Trust Distribution by Region (Scale: 1–4)"
#| fig-align: center
#| echo: false
#| warning: false
#| message: false


p1 <- ggplot(df_clean, aes(x = Inst_Trust_Index, fill = Region)) +
  geom_histogram(
    aes(y = after_stat(count / sum(count))),
    binwidth = 0.25,
    color = "white",
    alpha = 0.7
  ) +
  facet_wrap(~Region, nrow = 1) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "Institutional Trust Distribution (Scale: 1–4)",
    x = "Trust Score (Higher = More Trust)",
    y = "Percentage"
  ) +
  theme(legend.position = "none")

p1
```

### **Methods**

We employ a four-stage modeling approach. All models use Ordinary Least Squares (OLS) regression with mean-centered independent variables (Institutional Trust) to ensure the interpretability of coefficients and interaction terms.

The first model establishes a global baseline across all regions. It evaluates the independent effects of perceived harm and institutional trust while controlling for regional "intercept shifts" and individual demographics. This model assumes the impact of harm and trust is identical across all societies.

$$\begin{aligned}
\text{ImmigrantsGood} = &\beta_0 + \beta_1(\text{InstitutionTrust}) \\
&+ \beta_2(\text{Region}) + \beta_3(\text{Edu}) + \beta_4(\text{Income}) + \beta_5(\text{Age}) + \epsilon
\end{aligned}$$

Model 2 estimates specific region regressions to allow all coefficients to vary freely across societies. This specification relaxes the pooled assumption in Model 1 and provides a descriptive comparison of how institutional trust and demographic factors relate to immigration attitudes within each regional context.

$$
\begin{aligned}
\text{ImmigrantsGood}_i =\ & \beta_0  + \beta_1 \text{InstitutionTrust}_i \\ & + \beta_2 \text{Edu}_i + \beta_3 \text{Income}_i + \beta_4 \text{Age}_i + \varepsilon_i \end{aligned}
$$


Model 3 introduces interaction terms between institutional trust and region indicators, allowing the marginal effect of institutional trust on immigration attitudes to vary across societies while retaining region-specific intercepts.

$$
\begin{aligned}
\text{ImmigrantsGood}_i =\ & \beta_0 + \beta_1 \text{InstitutionTrust}_i + \boldsymbol{\beta_2}\,\text{Region}_i \\ &+ \boldsymbol{\beta_3}\, (\text{InstitutionTrust}_i \times \text{Region}_i) + \beta_4 \text{Edu}_i + \beta_5 \text{Income}_i + \beta_6 \text{Age}_i + \varepsilon_i \end{aligned}
$$

### **Results**

Table 1 shows the results for the simplest formula, Institutional trust is positively and statistically significantly associated with more favorable perceptions of immigration’s impact on the country. On average, a one-unit increase in institutional trust is associated with a 0.25 point increase in the perceived impact score. Relative to the reference region (Hong Kong), respondents in Singapore, Taiwan, and China report significantly more positive baseline attitudes toward immigration, with the largest intercept difference observed in China. Education and income are also positively related to immigration attitudes, while age has no statistically significant association. Overall, the model explains approximately 15 percent of the variation in immigration perceptions, providing a baseline estimate of the average relationship between institutional trust and immigration attitudes across regions.

```{r}
#| echo: false
#| warning: false
#| message: false
#| results: asis
stargazer(model_simpliest, 
          type = "latex",
          title = "Model 1 - Pooled Additive Model Results",
          column.labels = "Full Sample",
          dep.var.labels = "Impact on Country",
          no.space = TRUE,
          single.row = TRUE,
          font.size = "small",
          digits = 2,
          header = FALSE, 
          keep.stat = c("n", "rsq", "adj.rsq"))
```

Table 2 presents the individual regional models, reconfirming the trends while introducing our variable of interest: the interaction term. Across all four societies, institutional trust is positively and significantly associated with more favorable views of immigration’s impact on the country, though the magnitude of this relationship varies substantially by region. The effect is strongest in Hong Kong (0.38) and Singapore (0.35), moderate in Taiwan (0.2), and substantially weaker in China (0.07). While education and income matter in some contexts, their effects are inconsistent across regions. Notably, the model explains very little variance in China, suggesting that individual-level characteristics provide limited explanatory power in that context.

```{r}
#| echo: false
#| warning: false
#| message: false
#| results: asis
stargazer(model_hkg, model_sgp, model_twn, model_chn,
          type = "latex",
          title = "Model 2 - Regional Interaction Model Results",
          column.labels = c("Hong Kong", "Singapore", "Taiwan", "China"),
          dep.var.labels = "Impact on Country",
          no.space = TRUE,
          single.row = TRUE,
          font.size = "small",
          digits = 2,
          header = FALSE,
          keep.stat = c("n", "rsq", "adj.rsq"))
```

Table 3 presents results from a pooled regression model with region trust interaction terms to test whether the association between institutional trust and perceptions of immigration’s impact varies across societies. Institutional trust is positively and strongly associated with more favorable views of immigration in the baseline region (Hong Kong). However, this relationship is significantly weaker in Taiwan and especially in China, as indicated by the negative and statistically significant interaction terms. In contrast, the interaction between institutional trust and Singapore is small and not statistically significant, suggesting that the effect of trust in Singapore is comparable to that in Hong Kong. Together, these results demonstrate substantial regional heterogeneity: while institutional trust consistently predicts more positive immigration attitudes, its marginal effect depends heavily on political and social context, with the weakest translation of trust into immigration support observed in China.

```{r}
#| echo: false
#| warning: false
#| message: false
#| results: asis
stargazer(model_full,
          type = "latex",
          title = "Model 3 - Full Triple-Interaction Model Results",
          column.labels = "Global Interaction",
          dep.var.labels = "Impact on Country",
          no.space = TRUE,
          single.row = TRUE,
          font.size = "small",
          digits = 2,
          header = FALSE,
          keep.stat = c("n", "rsq", "adj.rsq"))
```

### **Discussion**

The results suggest that trust in government institutions is a consistent and robust predictor of immigration attitudes across the sampled East Asian societies. It aligns with our initial hypothesis, individuals who perceive immigrants as posing greater threats to crime, security, or the economy are less likely to believe that immigration benefits national development. Conversely, higher levels of institutional trust, which are measured through confidence in the police, courts, and civil service, are associated with more favorable evaluations of immigration’s impact.
At the same time, the pooled results indicate notable regional differences in baseline attitudes. Relative to Hong Kong, respondents in China, Singapore, and Taiwan exhibit more positive baseline perceptions of immigration. This pattern suggests that Hong Kong’s distinct socio-political context may foster a more skeptical starting point for immigration attitudes, potentially reflecting localized political tensions and social dynamics rather than generalized opposition to immigration per se.

Our findings are consistent with evidence from other political contexts. Prior research on the American mass public demonstrates that individuals with higher levels of political trust are more supportive of pro-immigration policies, even after accounting for partisanship, attitudes toward immigrant groups, and core political values. Together, these results suggest that the positive association between institutional trust and immigration attitudes is not confined to Western democracies but may operate across diverse political and cultural settings. In both the United States and East Asia, higher trust in political institutions appears to correspond with more welcoming views of immigration, underscoring the broader political relevance of institutional trust as a key determinant of immigration attitudes.

Nevertheless, the analysis is subject to potential endogeneity concerns. Unobserved predispositions, such as nationalism or generalized hostility toward out-groups, may simultaneously shape perceptions of immigrant-related harm and trust in political institutions. As a result, while the observed associations are robust, the findings should be interpreted as correlational rather than strictly causal.


### **Reference**

“List of Sovereign States by Immigrant and Emigrant Population.” Wikipedia, 10 Mar. 2024, en.wikipedia.org/wiki/List_of_sovereign_states_by_immigrant_and_emigrant_population.

Hainmueller, J., & Hopkins, D. J. (2014). Public attitudes toward immigration. Annual review of political science, 17(1), 225-249.

Stoyan, A. T., & Niedzwiecki, S. (2018). Decentralization and democratic participation: The effect of subnational self-rule on voting in Latin America and the Caribbean. Electoral Studies, 52, 26–35.[ ](https://doi.org/10.1016/j.electstud.2017.12.001)

Heo, N. (2024). “Generalized trust and attitudes toward immigrants in East Asia.” International Journal of Intercultural Relations.

Macdonald, David. “Political Trust and Support for Immigration in the American Mass Public.” British Journal of Political Science, vol. 51, no. 4, 17 Mar. 2020, pp. 1–19, https://doi.org/10.1017/s0007123419000668.
